Capstone Movielens Report

### 2025-06-03 reports/capstone-movielens.site/02.6.UMGE-model.Rmd
Started working on new model: model template added.
MF: Mathematical Description section completed.

To account for a _genre effect_ we will use the model suggested in the [Section 23.7: Exercises](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#exercises) of the _Chapter "23 Regularization" of the Course Textbook_[@IDS2_23-7]:

@@

\begin{equation}
\hat{g}_{i,j}(\lambda) = \frac{1}{\lambda + n_g} \sum_{i=1}^{n_{g}} \left(Y_{i,j} - \mu - \alpha_i - \beta_j\right)
(\#eq:GE-regularized)
\end{equation}

>>>>

To account for a _genre effect_ we will use the model suggested in the [Section 23.7: Exercises](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#exercises) of the _Chapter 23 "Regularization" of the Course Textbook_[@IDS2_23-7]:

@@

\begin{equation}
\hat{g}_{i,j}(\lambda) = \frac{1}{\lambda + n_g} \sum_{r=1}^{n_{g}} \left(Y_{i,j} - \mu - \alpha_i - \beta_j\right)
(\#eq:GE-regularized)
\end{equation}

### 2025-06-03 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
MF: Mathematical Description section completed.

#### UMGYDE Model: Final Holdout Test (Preliminary Assesment)

@@

final.RMSEs.ResultTibble.UMGYDE.rglr.tuned <- RMSEs.ResultTibble.UMGYDE.rglr.tuned |> 
  RMSEs.AddRow("Best UMGYDE Model (Final Test)", 
               final.UMGYDE.predicted.RMSE,
               comment = "Final Holdout Test of the best tuned and regularized UMGYDE Model.")
```
```{r}
RMSE_kable(final.RMSEs.ResultTibble.UMGYDE.rglr.tuned)
```

@@

@@

>>>>

#### UMGYDE Model: Final Holdout Test (Preliminary Assesment)
\

@@

final.RMSEs.ResultTibble.UMGYDE.rglr.tuned <- RMSEs.ResultTibble.UMGYDE.rglr.tuned |> 
  RMSEs.AddRow("Best UMGYDE Model (Final Test)", 
               final.UMGYDE.predicted.RMSE,
               comment = "Final Holdout Test of the best tuned and regularized UMGYDE Model.")
```

@@

And finally, we print out the results:
```{r}
RMSE_kable(final.RMSEs.ResultTibble.UMGYDE.rglr.tuned)
```

@@

As explained in the [Chapter 24 `Matrix Factorization`](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/matrix-factorization.html) of the _Course Textbook_, so far our models _"ignore an important source of information related to the fact that groups of movies have similar rating patterns and groups of users have similar rating patterns as well..."_ [@IDS2_24]

Then the author shows that in this case, the _Matrix Factorization_ method can considerably improve our results. In the next section, we will apply this method to our current model and see what we get.

### 2025-06-03 reports/capstone-movielens.site/02.8-1.MF.UMGYDE-model.Rmd
MF: Mathematical Description section completed.

### UMGYDE Model: Matrix Factorization
\

@@

>>>>

## UMGYDE Model: Matrix Factorization (MF)
\

For our next solution, we will use the [`recosystem` package](https://cran.r-project.org/web/packages/recosystem/index.html) to perform the _Parallel Matrix Factorization_.

### MF: Matematical Description
\

As outlined in [this article](https://cran.r-project.org/web/packages/recosystem/vignettes/introduction.html)[@yqiu_recosystem230505], the idea of the _Matrix Factorization_ method is to approximate the whole rating matrix $R_{m \times n}$ by the product of two matrices of lower dimensions, $P_{n \times k}$ and $Q_{n\times k}$, such that

\begin{equation}
\mathbb{R} \approx \mathbb{PQ^T}
(\#eq:R-PQ-MF)
\end{equation}

In relation to our model, the expression \@ref(eq:R-PQ-MF) will take the form:

\begin{equation}
\mathbf{R} \thicksim \mathbf{\hat{R}} + \mathbf{PQ^T} +  \varepsilon
(\#eq:UMGYDE-MF)
\end{equation}

where:

  - $\mathbf{R}$ is the $U_\mathrm{m} \times M_n$ rating matrix with $U_m$ users and $M_n$ movies;
  - $\mathbf{\hat{R}}$ represents the predictions from our best model: _Regularized UMGYDE Model_;
  - $\mathbf{P}$ and $\mathbf{Q}$ are $P_{m \times k}$ and $Q_{n \times k}$ matrices, respectively, where $k$ is the number of _latent features_ to be found.

If we denote the $u$-th row of $\mathbf{P}$ as $\mathbf{p}_u$ and the $v$-th row of $\mathbf{Q}$ as $\mathbf{q}_v$, then the unknown rating $\mathbf{r}_{u,v}$ given by user $u$ on movie item $v$ for our model can be estimated as $\hat{\mathbf{r}}_{u,v} + \mathbf{p}_{u}\mathbf{q}^T_{v}$, where the $\hat{\mathbf{r}}_{u,v}$ is the prediction given by our last _Regularized UMGYDE Model_.

A typical solution for $\mathbf{P}$ and $\mathbf{Q}$ is given by the following optimization problem [@chin_FPSGM_MF_SMS2015a; @chin_LRS_SGM_MF2015b]:

$$
\begin{array}{cc} min \\
\tiny{P,Q} \end{array} \sum_{(u,v)\in R}[f(\mathbf{p}_u, \mathbf{q}_v; \mathbf{r}_{u,v}) + \mu_P\parallel\mathbf{p}_u\parallel_1 + \mu_Q\parallel\mathbf{q}_v\parallel_1 + \frac{\lambda_P}2\parallel\mathbf{p}_u\parallel^2_2 + \frac{\lambda_Q}2\parallel\mathbf{q}_v\parallel^2_2]
$$

where $(u, v)$ are locations of observed entries in $\mathbf{R}, \mathbf{r}_{u,v}$ is the observed ratings, $f$ is the loss function, and $\mu_P$, $\mu_Q$, $\lambda_P$, $\lambda_Q$ are penalty parameters to avoid overfitting.[@yqiu_recosystem230505]

::: { custom-style="blockquote" } 
> _The process of solving the matrices $P$ and $Q$ is referred to as model training, and the selection of penalty parameters is called parameter tuning. In `recosystem`, we provide convenient functions for these two tasks, and additionally have functions for model exporting (outputting $P$ and $Q$ matrices) and prediction._[@yqiu_recosystem230505]

:::

@@

### MF: Helper Funtions
